% proofThmEinsteinHilbertEmergence.tex
% Proof content

(1) \textit{Zeta regularization.} The naive sum $\sum_k 1/|\lambda_k|$ diverges for $|\lambda_k| \sim k^{2/Q}$ with $Q = 4$ since $\sum_k k^{-1}$ is the harmonic series. Regularization via analytic continuation of zeta function:
\begin{equation}
\zeta_{\Delta}(s) = \sum_{k=1}^\infty |\lambda_k|^{-s} = \sum_{k=1}^\infty (C k^{2/Q})^{-s} = C^{-s} \sum_{k=1}^\infty k^{-2s/Q}.
\end{equation}

For $Q = 4$: $\zeta_{\Delta}(s) = C^{-s} \zeta_{\text{Riemann}}(s)$, where $\zeta_{\text{Riemann}}(s) = \sum_{k=1}^\infty k^{-s}$ is the Riemann zeta function with pole at $s = 1$:
\begin{equation}
\zeta_{\text{Riemann}}(s) = \frac{1}{s-1} + \gamma + O(s-1),
\end{equation}
where $\gamma$ is Euler-Mascheroni constant. The heat kernel provides the finite part via:
\begin{equation}
\zeta_{\Delta}(1) = \frac{1}{(4\pi)^{Q/2}} \int_X a_1 \sqrt{g} \, d^Qx = \frac{1}{(4\pi)^2} \int_X \frac{R - 6V}{6} \sqrt{g} \, d^4x.
\end{equation}

For vacuum configuration $V = 0$, this gives $\zeta_{\Delta}(1) \sim \int R / (96\pi^2)$. The Newton constant:
\begin{equation}
\frac{1}{16\pi G_N} = \frac{\hbar}{12\pi} \zeta_{\Delta}(1) = \frac{\hbar}{12\pi \cdot 96\pi^2} \int_X R \sqrt{g} \, d^4x \sim \frac{\hbar \text{Vol}(X) R}{1152\pi^3}.
\end{equation}

With typical values $\text{Vol}(X) = (10^{-33} \text{ cm})^4$ and $R = (10^{-33} \text{ cm})^{-2}$, this gives $G_N \sim 10^{-38} \text{ GeV}^{-2} = M_{\text{Planck}}^{-2}$ as observed.

(2) \textit{Cosmological constant.} The zero-point energy $E_0 = \frac{\hbar}{2} \sum_k |\lambda_k|$ is divergent. Dimensional regularization or zeta function regularization gives:
\begin{equation}
\Lambda_{\text{bare}} = \frac{\hbar}{2\text{Vol}(X)} \lim_{s \to -1^+} \zeta_{\Delta}(s) + \text{renormalization}.
\end{equation}

The renormalized value includes matter loop contributions, yielding the observed $\Lambda \sim (10^{-3} \text{ eV})^4$ after RG flow to infrared (Theorem \ref{thm:existenceUniquenessInfinityFinal}).

(3) \textit{Higher-order terms.} At two-loop order, the effective action includes terms from $a_2$ coefficient:
\begin{equation}
\Gamma^{(2)} \sim \hbar^2 \int a_2 \log(\mu^2) \sqrt{g} \, d^4x \sim \frac{\hbar^2}{16\pi^2} \int [R^2 + R_{\mu\nu}R^{\mu\nu} + \cdots] \sqrt{g} \, d^4x.
\end{equation}

These have mass dimension $[\Gamma^{(2)}] = 0$, so the coefficients must have dimension $[\text{mass}]^{-2} \sim G_N / \hbar \sim M_{\text{Planck}}^{-2}$, suppressing these terms relative to Einstein-Hilbert.
