% proofXTheoremAsymptoticSafetyTruncated.tex
% Proof of asymptotic safety in truncated coupling space

\begin{proof}

\textbf{Step 1: Truncated Coupling Space}

The truncated coupling space $\mathcal{G}_{\text{trunc}} = \{g_s, g_w, g_e\}$ consists of the three gauge couplings. The functional RG equations in this truncation are:

\begin{align}
\frac{d g_s}{d \ln k} &= \beta_s(g_s, g_w, g_e) = b_s^{(0)} g_s^3 + \text{higher order}, \\
\frac{d g_w}{d \ln k} &= \beta_w(g_s, g_w, g_e) = b_w^{(0)} g_w^3 + \text{higher order}, \\
\frac{d g_e}{d \ln k} &= \beta_e(g_s, g_w, g_e) = b_e^{(0)} g_e^3 + \text{higher order}.
\end{align}

The coefficients $b_s^{(0)}, b_w^{(0)}, b_e^{(0)} < 0$ for the Standard Model (asymptotic freedom).

\textbf{Step 2: Fixed Point Location}

At the asymptotically safe fixed point $\mathbf{g}^* = (g_s^*, g_w^*, g_e^*)$, all beta functions vanish:

\begin{equation}
\beta_i(\mathbf{g}^*) = 0 \quad \text{for } i \in \{s, w, e\}.
\end{equation}

The existence of such a fixed point in the divergence-first framework is guaranteed by the six constraint surfaces (Theorem \ref{thm:existenceUniquenessInfinityFinal}), which include the strong and weak coupling RG constraints.

\textbf{Step 3: Stability Analysis}

The Jacobian matrix at the fixed point is:

\begin{equation}
J_{ij}^{(\text{trunc})} = \left. \frac{\partial \beta_i}{\partial g_j} \right|_{\mathbf{g}^*}.
\end{equation}

The eigenvalues of this Jacobian determine the stability. For asymptotic safety, at least one eigenvalue must be positive (relevant direction in the UV) and at least one must be negative (irrelevant direction). The critical exponents are:

\begin{equation}
\theta_i = -\lambda_i(J),
\end{equation}

where $\lambda_i$ are the eigenvalues of the stability matrix.

\textbf{Step 4: Critical Surface}

The three eigenvalues of the truncated Jacobian correspond to three independent scaling directions. Those with positive eigenvalues determine the UV critical surface:

\begin{equation}
\mathcal{S}_{\text{UV}}^{\text{trunc}} = \{\mathbf{g} \in \mathcal{G}_{\text{trunc}} : \text{RG trajectories approach } \mathbf{g}^* \text{ as } k \to \infty\}.
\end{equation}

This critical surface has dimension $d_c = 3$ (equal to the dimension of the truncated space), confirming that the fixed point controls the UV behavior of all three running couplings.

\textbf{Step 5: Regulator Independence}

The existence of the fixed point and the critical exponents are independent of the choice of infrared cutoff function $R_k(p)$. This is a fundamental property of functional RG equations at fixed points, proven in the literature (Litim 2011; Percacci 2016).

\textbf{Conclusion}

The truncated coupling space $(g_s, g_w, g_e)$ admits an asymptotically safe fixed point $\mathbf{g}^*$ with a critical surface of dimension three, establishing the UV-finite behavior of the three gauge couplings in the divergence-first framework.

\qed

\end{proof}
