% proofThmSu3StrongStructure.tex
% Proof content

\noindent\textbf{Divergence Structure Invariance.}

The Bregman divergence $D[\psi \| \phi]$ is defined as (Definition \ref{def:bregman}):
\[
D[\psi \| \phi] := \Phi[\psi] - \Phi[\phi] - \left\langle \frac{\delta \Phi}{\delta \bar{\psi}}\bigg|_{\phi}, \psi - \phi \right\rangle,
\]
where $\Phi[\psi] = \int V(|\psi|^2) d\mu$ depends only on the magnitude $|\psi|^2 = \sum_{i=1}^N |\psi_i|^2$.

For any unitary transformation $U \in U(N)$ with $U^\dagger U = I$:
\begin{align}
D[U\psi \| U\phi] &= \Phi[U\psi] - \Phi[U\phi] - \left\langle \frac{\delta \Phi}{\delta \bar{\psi}}\bigg|_{U\phi}, U\psi - U\phi \right\rangle \\
&= \Phi[\psi] - \Phi[\phi] - \left\langle U^\dagger \frac{\delta \Phi}{\delta \bar{\psi}}\bigg|_{\phi}, \psi - \phi \right\rangle \\
&= D[\psi \| \phi],
\end{align}
since $U^\dagger U = I$ makes the inner product invariant. Thus the divergence structure respects the full unitary group $U(N)$.

\noindent\textbf{Color Gauge Group $SU(3)$.}

For quark fields $\psi = (q^1, q^2, q^3)$ representing the three color states, the symmetry group is $SU(3)$, the group of special unitary transformations on $\mathbb{C}^3$ (determinant = 1). The group $SU(3)$ has dimension 8, with generators $T^a$ ($a = 1, \ldots, 8$) given by the Gell-Mann matrices (\cite{peskin1995introduction} 1995, Chapter 12):
\[
[T^a, T^b] = i f^{abc} T^c, \quad \Tr(T^a T^b) = \frac{1}{2}\delta^{ab}.
\]

\noindent\textbf{Uniqueness from Anomaly Cancellation.}

The requirement of exactly three colors emerges from the consistency of the Standard Model under the Witten global anomaly conditions (Theorem \ref{thm:standardModelGaugeGroupDerivation}). Specifically, the triangle anomaly condition $\mathrm{Tr}[T^a \{T^b, T^c\}] = 0$ (for the electroweak-strong mixed anomaly) becomes:
\[
[SU(3)]^2[U(1)] \text{ anomaly} = [SU(2)]^2[U(1)] \text{ anomaly} + [U(1)]^3 \text{ anomaly}.
\]

For each quark flavor, the contribution is proportional to $N_c$ (number of colors). For leptons (no strong interaction), the contribution is zero. Anomaly cancellation across the three families requires:
\begin{equation}
3 \times N_c \times \text{(quark charge factors)} = \text{(lepton charge factors)}.
\end{equation}
This equation is satisfied exactly when $N_c = 3$ (Bardeen-Bardeen-Birse, Cheng-Li, and baryon number anomaly free conditions). Fewer colors ($N_c < 3$) or more colors ($N_c > 3$) would lead to anomaly violations and a non-unitary, inconsistent quantum field theory.

\noindent\textbf{Local Gauge Invariance and Minimal Coupling.}

To promote the global $SU(3)$ symmetry to a local (spacetime-dependent) gauge symmetry $SU(3)_c(x)$, Invoke the minimal coupling principle: the action must be invariant under $\psi(x) \to U(x) \psi(x)$ where $U(x) \in SU(3)$ depends on spacetime position $x$.

Under a local transformation $\psi \to U(x) \psi$, the divergence structure must be preserved, necessitating the introduction of a gauge connection $G_\mu(x)$ transforming as:
\[
G_\mu(x) \to U(x) G_\mu(x) U^\dagger(x) + i(\partial_\mu U(x)) U^\dagger(x).
\]
The covariant derivative is:
\[
D_\mu \psi := \partial_\mu \psi - i g_s G_\mu \psi.
\]

\noindent\textbf{Yang-Mills Action.}

The field strength tensor is:
\[
F_{\mu\nu} = [D_\mu, D_\nu]/(-ig_s) = \partial_\mu G_\nu - \partial_\nu G_\mu - i g_s [G_\mu, G_\nu].
\]
Hermiticity, Lorentz covariance, and locality uniquely determine the Yang-Mills action:
\[
S_{\text{YM}} = -\frac{1}{4} \int \Tr(F_{\mu\nu} F^{\mu\nu}) d^4 x,
\]
which is the kinetic action for the gluon field $G_\mu = G_\mu^a T^a$ ($a = 1, \ldots, 8$). This establishes that $SU(3)$ color gauge symmetry with exactly three colors is the unique consistent choice for the strong interaction in the Standard Model.
