% Part of Section 0: Introduction
\subsection{Minimal Axiomatic Structure}
\label{subsec:axioms}

The framework rests on two axioms formulated with maximum economy. These suffice to generate all of fundamental physics. All additional structures are imposed.

The first axiom specifies a metric measure space: a topological space equipped with both distance and a probability measure. Two geometric regularity conditions apply. Ahlfors regularity requires that measure scales uniformly as a power law function of distance. A Poincaré inequality constrains the oscillations of functions relative to both distance and measure. These conditions ensure well-behaved functional analysis on the space without constraining dimensionality. The dimension parameter is left entirely unspecified at the axiom level and determined later through consistency requirements.

The second axiom specifies that the configuration space consists of square-integrable complex-valued functions over the metric measure space. These functions are governed by a strictly convex generating functional. The strict convexity property is fundamental: it ensures that unique extremal configurations exist and that polarization of the functional yields a well-defined quadratic form. All assumptions about particles, fields, forces, or gauge structures enter at the axiom level.

From the strict convexity of the generating functional emerges an asymmetric Bregman divergence. This divergence is not imposed as an additional axiom but derived from the functional structure. The Hessian of the generating functional yields three independent eigenvalue scales corresponding to soft, bulk, and stiff modes. These three information channels are fundamental consequences of the axiomatization and determine that exactly three fermion generations must exist through representation-theoretic structure. This ternary organization is derived, not assumed. The internal structure emerges from mathematical necessity rather than external specification.

All spacetime geometry, quantum mechanics, forces, and particle content follow from these two axioms through rigorous logical derivation. Nothing more is assumed.
