% Part of Section 0: Introduction
\subsection{The Ordered Emergence Sequence}
\label{subsec:emergenceLogic}

The mathematical development follows these axioms in a strictly ordered logical sequence. Each result depends exclusively on previously established results. All circular reasoning occurs and the logical dependency structure has been verified acyclic through topological sorting.

From the convexity structure specified in the second axiom emerges a symmetric quadratic form through polarization of the divergence functional. This form defines a Dirichlet form, which is a self-adjoint differential structure governing diffusion and harmonic functions. The Dirichlet form representation theorem determines a unique self-adjoint spectral operator. The eigenfunctions of this operator form an orthonormal basis for the function space.

These eigenfunctions must satisfy regularity conditions that link their oscillation properties to the dimensionality of the underlying space. Hölder continuity of eigenfunctions is a mathematical requirement that bounded oscillations exist. This condition forces the Ahlfors scaling dimension strictly below four. For dimension equal to or greater than four, no continuous eigenfunction can exist. Mathematical consistency eliminates all higher-dimensional possibilities.

From the spectral structure emerges a Riemannian metric through the Carré du Champ operator formalism. This classical construction uses eigenfunction oscillations to define distance relationships. Spectral embedding constructs smooth manifold geometry where the original space embeds into Euclidean space with distances preserved. The resulting structure is a smooth manifold with well-defined tangent spaces and smooth transition functions.

Asymmetry inherent in Bregman divergence breaks time reversal symmetry and induces causal ordering on the manifold. The Osterwalder-Schrader reconstruction theorem performs analytic continuation from Euclidean metric geometry to Lorentzian spacetime. The continuation yields one temporal and three spatial directions. Lorentzian signature emerges as a forced mathematical consequence of divergence asymmetry applied to geometry emergence.

Four-dimensional spacetime with one temporal and three spatial directions has emerged from measure-theoretic axioms alone. All assumption about dimensionality or signature has entered the framework. Geometry is generated rather than assumed.
