% sectionOEffectiveActionGravity.tex
% Section content



\section{Effective Action and Emergent Gravity}
\label{sec:effectiveActionGravity}
\label{sec:effectiveAction}


\input{epigraphRiemann}

\subsection{One-Loop Effective Action from Heat Kernel}
\label{subsec:oneLoopEffectiveActionFromHeatKernel}

\begin{definition}[Effective Average Action with Quantum Corrections]
\label{def:effectiveAction}
The one-loop effective action:
\begin{equation}
\Gamma[g] := S_{\text{classical}}[g] + \frac{\hbar}{2} \Tr \log(H_g) + O(\hbar^2),
\end{equation}
where $H_g := -\nabla_g^2 + V''(\psi_{\text{cl}})$ is the Hessian at the classical solution, and $\Tr \log$ is regularized via heat kernel methods.
\end{definition}

\begin{theorem}[Heat Kernel Expansion and Seeley-DeWitt Coefficients]
\label{thm:seeleyDewitt}
For a differential operator $P = -\nabla^2 + V$ on compact Riemannian $(X, g)$, the heat kernel $K_t(x, y)$ satisfies $(\partial_t + P_x) K_t(x, y) = 0$ with $K_0(x, y) = \delta(x - y)$.

The trace expansion:
\begin{equation}
\Tr K_t := \int_X K_t(x, x) \sqrt{g} \, d^4x \sim (4\pi t)^{-2} \sum_{n=0}^\infty a_n t^n
\end{equation}
has Seeley-DeWitt coefficients:

\begin{enumerate}
\item $a_0 = \int_X \sqrt{g} \, d^4x = \text{Vol}(X)$.

\item $a_1 = \frac{1}{6} \int_X (R - 6V) \sqrt{g} \, d^4x$.

\item $a_2 = \frac{1}{360} \int_X [12R^2 - 5R_{\mu\nu}R^{\mu\nu} + 2R_{\mu\nu\rho\sigma}R^{\mu\nu\rho\sigma} + 30V^2 - 60RV] \sqrt{g} \, d^4x$.
\end{enumerate}

The functional determinant:
\begin{equation}
\Tr \log H_g = \frac{a_0}{2\epsilon^2} - \frac{a_1}{2\epsilon} + a_2 \log(\mu^2) + \text{finite},
\end{equation}
where $\epsilon$ is a UV cutoff and $\mu$ is the renormalization scale.

\begin{proof}
\input{proofOTheoremSeeleyDewitt}
\end{proof}
\end{theorem}

\subsection{IR Limit and Effective Action Extraction}
\label{subsec:irLimitEffectiveAction}

\begin{lemma}[IR Limit Commutes with Functional Integration]
\label{lem:irLimitCommutes}

For the one-loop effective action in the divergence-first framework:
\begin{equation}
S_{\text{eff}}[g; k] := S_0[g] + \frac{1}{2} \Tr \log\left(\Gamma_k^{(2)}[g] + R_k\right),
\end{equation}

the infrared limit $k \to 0$ can be interchanged with the functional integral over metric fields. The proof uses scale separation and monotonicity of the RG flow.

\begin{proof}
\input{proofOLemmaIrLimitCommutes}
\end{proof}
\end{lemma}

\begin{theorem}[Einstein-Hilbert Action Emerges from One-Loop Effective Action]
\label{thm:einsteinHilbertEmergence}
Computing the one-loop effective action from Seeley-DeWitt coefficients:
\begin{equation}
\Gamma_{\text{eff}}[g] = \frac{1}{16\pi G_N} \int_X (R - 2\Lambda) \sqrt{g} \, d^4x + \int_X \mathcal{L}_{\text{matter}}(g, \psi) \sqrt{g} \, d^4x + O(R^2),
\end{equation}

where:

\begin{enumerate}
\item \textbf{Newton Constant from Spectral Zeta Regularization.} The gravitational constant emerges from the regularized spectral sum:
\begin{equation}
\frac{1}{16\pi G_N} := \frac{\hbar}{12\pi} \lim_{s \to 1^+} \left[\zeta_{\Delta}(s) - \frac{Q/2}{s-1}\right],
\end{equation}
where $\zeta_{\Delta}(s) := \sum_{k=1}^\infty |\lambda_k|^{-s}$ is the spectral zeta function of $-\Delta_\mu$.

For $Q = 4$: The zeta function $\zeta_{\Delta}(s) = \sum_k k^{-2s/Q} = \sum_k k^{-s}$ converges for $\text{Re}(s) > 2$ and admits meromorphic continuation to $s = 1$ with simple pole:
\begin{equation}
\zeta_{\Delta}(s) = \frac{Q/2}{s-1} + \zeta_{\Delta}(1) + O(s-1),
\end{equation}
where $\zeta_{\Delta}(1)$ is finite and given by heat kernel asymptotics:
\begin{equation}
\zeta_{\Delta}(1) = \frac{1}{(4\pi)^2} \int_X \frac{R}{6} \sqrt{g} \, d^4x
\end{equation}
from Seeley-DeWitt coefficient $a_1$.

Thus:
\begin{equation}
\frac{1}{16\pi G_N} = \frac{\hbar}{12\pi} \cdot \frac{1}{(4\pi)^2} \int_X \frac{R}{6} \sqrt{g} \, d^4x = \frac{\hbar}{12\pi} \cdot \frac{\text{Vol}(X) R}{96\pi^2}.
\end{equation}

Setting $\text{Vol}(X) \sim \ell_{\text{Planck}}^4$ and $R \sim \ell_{\text{Planck}}^{-2}$ (curvature at Planck scale):
\begin{equation}
G_N \sim \frac{12\pi \cdot 96\pi^2}{\hbar \ell_{\text{Planck}}^2} \sim \frac{\ell_{\text{Planck}}^2}{\hbar} \sim M_{\text{Planck}}^{-2},
\end{equation}
recovering correct dimensionality $[G_N] = [\text{mass}]^{-2}$ in natural units.

\item \textbf{Cosmological Constant from Zero-Point Energy.} The cosmological constant arises from vacuum energy:
\begin{equation}
\Lambda := \frac{\hbar}{2 \text{Vol}(X)} \left[\sum_{k=1}^\infty |\lambda_k| - \int_0^\infty \frac{d\lambda}{\lambda} \rho(\lambda)\right] + \Delta_{\text{matter}},
\end{equation}
where $\rho(\lambda)$ is the density of states and the integral provides the continuum subtraction (removing divergent constant). The finite remainder plus matter contributions $\Delta_{\text{matter}}$ determines the observed $\Lambda$.

\item \textbf{Higher-Curvature Terms.} Appear at two-loop order from Seeley-DeWitt coefficient $a_2$, giving Weyl-squared and Gauss-Bonnet terms:
\begin{equation}
\Gamma^{(2)}[g] \sim \frac{\hbar^2}{M_{\text{Planck}}^2} \int_X [c_1 R^2 + c_2 R_{\mu\nu}R^{\mu\nu} + c_3 R_{\mu\nu\rho\sigma}R^{\mu\nu\rho\sigma}] \sqrt{g} \, d^4x,
\end{equation}
suppressed by $1/M_{\text{Planck}}^2$ relative to Einstein-Hilbert term.
\end{enumerate}

Thus Einstein-Hilbert gravity emerges as the low-energy effective theory.

\begin{proof}
\input{proofOTheoremEinsteinHilbertEmergence}
\end{proof}
\end{theorem}

\begin{theorem}[Gravitational Scaling Limit and IR Effective Action]
\label{thm:scalingLimitGravity}

Define the gravitational scaling limit on a family of scaled metric measure spaces $(X_\ell, d_\ell, \mu_\ell)$ parameterized by length scale $\ell > 0$:
\begin{equation}
X_\ell := \ell X, \quad d_\ell(x, y) := \ell \cdot d_X(x/\ell, y/\ell), \quad \mu_\ell(E) := \ell^{-Q} \mu(E/\ell).
\end{equation}

The Laplacian on the scaled space is $-\Delta_\ell := \ell^{-2} (-\Delta)$, so eigenvalues scale as $\lambda_k^\ell := \ell^{-2} \lambda_k$.

\begin{enumerate}[label=(\roman*)]

\item \textbf{Infrared Limit.} As $\ell \to \infty$ (infrared limit), the spectrum of $-\Delta_\ell$ becomes denser, approaching a continuum. The density of states becomes:
\begin{equation}
\rho(\lambda) \sim \frac{\text{Vol}(X)}{(4\pi)^{Q/2}} \lambda^{Q/2 - 1},
\end{equation}
which by Weyl asymptotics (Theorem \ref{thm:WeylAsymptotics}) for $Q = 4$ is $\rho(\lambda) \sim \frac{\text{Vol}(X)}{(4\pi)^2} \lambda$.

\item \textbf{One-Loop Effective Action in IR Limit.} The one-loop effective action in this limit is:
\begin{equation}
\Gamma_{\text{IR}}[g] := \lim_{\ell \to \infty} \Gamma[\ell^{-2}g],
\end{equation}
where the metric rescaling $\ell^{-2}g$ accounts for the spectral scaling. This IR effective action takes the form:
\begin{equation}
\Gamma_{\text{IR}}[g] = \int_X \left[\frac{1}{16\pi G_N} (R - 2\Lambda) + \mathcal{L}_{\text{matter}}\right] \sqrt{g} \, d^4x + O(\hbar^2),
\end{equation}
with Newton constant emerging from spectral regularization.

\item \textbf{Planck Scale from Spectral Gap.} The Planck length arises naturally as:
\begin{equation}
\ell_P := \sqrt{\hbar G_N} \sim \ell_0 \epsilon^{1/2},
\end{equation}
where $\ell_0$ is the microscopic length scale (set by the minimum spectral gap $|\lambda_1|$) and $\epsilon$ is a dimensionless coupling constant.

\item \textbf{Renormalization Group Flow to Einstein Gravity.} The IR effective action $\Gamma_{\text{IR}}[g]$ is a fixed point of the renormalization group flow and is independent of the UV cutoff (renormalization-group invariance). This ensures that Einstein gravity emerges as the unique universal low-energy effective theory from the microscopic spectral structure.

\end{enumerate}

\begin{proof}
\input{proofOTheoremScalingLimitGravity}
\end{proof}

\end{theorem}

\subsection{Gravitational Interaction as Nonlocal Entanglement}
\label{subsec:gravitationalInteractionAsNonlocalEntanglement}

\begin{theorem}[Gravity Emerges from Information Geometry]
\label{thm:gravitationalConvolution}
Gravitational attraction arises as nonlocal entanglement of mass distributions:
\begin{equation}
F(x_1, x_2) = \iint dx \, dy \, \frac{m_1(x)}{|x - x_1|} \cdot \frac{m_2(y)}{|y - x_2|},
\end{equation}
where $m_i(x) := |\psi_i(x)|^2$ are informational densities.

In weak-field limit, this bilinear convolution structure yields Newton's inverse-square law:
\begin{equation}
F(x_1, x_2) \approx \frac{M_1 M_2}{|x_1 - x_2|^2}
\end{equation}
where $M_i := \int m_i(x) dx$ are total masses.

\begin{proof}
\input{proofOTheoremGravitationalConvolution}
\end{proof}
\end{theorem}

\subsection{Two-Loop Effective Action and UV Finiteness Verification}
\label{subsec:twoLoopUVFiniteness}

\begin{theorem}[Two-Loop Effective Action Structure]
\label{thm:twoLoopEffectiveAction}

in the divergence-first framework with one-loop quantum corrections giving the Einstein-Hilbert action, the two-loop effective action contributes higher-derivative terms:

\begin{equation}
\Gamma^{(2)}[g] = \frac{\hbar^2}{M_{\text{Planck}}^2} \int_X [c_1 R^2 + c_2 R_{\mu\nu}R^{\mu\nu} + c_3 R_{\mu\nu\rho\sigma}R^{\mu\nu\rho\sigma}] \sqrt{g} \, d^4x,
\end{equation}

where the coefficients are:

\begin{enumerate}
\item $c_1 = \frac{1}{(4\pi)^2} \cdot \frac{1}{30}$ (from Weyl tensor contribution)
\item $c_2 = -\frac{1}{(4\pi)^2} \cdot \frac{1}{72}$ (from Ricci tensor contribution)
\item $c_3 = \frac{1}{(4\pi)^2} \cdot \frac{1}{180}$ (from Riemann tensor contribution)
\end{enumerate}

These arise from the Seeley-DeWitt coefficient $a_2$ and depend only on the topology and dimension $d = 4$.

\begin{proof}

\textbf{Part 1: Seeley-DeWitt Coefficient $a_2$ Computation}

From heat kernel asymptotics (Theorem \ref{thm:seeleyDewitt}), coefficient $a_2$ for a scalar Laplacian on Riemannian 4-manifold is:

\begin{equation}
a_2 = \frac{1}{360} \int_X [12R^2 - 5R_{\mu\nu}R^{\mu\nu} + 2R_{\mu\nu\rho\sigma}R^{\mu\nu\rho\sigma} + 30V^2 - 60RV] \sqrt{g} \, d^4x.
\end{equation}

This is the standard result from \cite{birrell1982quantum} and \cite{vassilevich2003heat}.

\textbf{Part 2: Integration of $a_2$ into Effective Action}

The two-loop effective action reads:

\begin{equation}
S_{\text{eff}}^{(2)}[g] = \frac{\hbar^2}{2} a_2[\phi_{\text{cl}}, g].
\end{equation}

Substituting and extracting pure geometric terms (setting $V \equiv 0$ at the field level, including only gravity-matter coupling):

\begin{equation}
\Gamma^{(2)}[g] = \frac{\hbar^2}{2 \cdot 360} \int_X [12R^2 - 5R_{\mu\nu}R^{\mu\nu} + 2R_{\mu\nu\rho\sigma}R^{\mu\nu\rho\sigma}] \sqrt{g} \, d^4x.
\end{equation}

Dimensionally, $\hbar^2$ has dimension [mass]$^{-2}$ in natural units. To match the dimension of the Ricci scalar $R$ (dimension [mass]$^2$), the must have explicit negative powers of $M_{\text{Planck}}$.

Thus:

\begin{equation}
\Gamma^{(2)}[g] = \frac{\hbar^2}{360 M_{\text{Planck}}^2} \int_X [12R^2 - 5R_{\mu\nu}R^{\mu\nu} + 2R_{\mu\nu\rho\sigma}R^{\mu\nu\rho\sigma}] \sqrt{g} \, d^4x,
\end{equation}

identifying:

\begin{align}
c_1 &= \frac{12}{360} = \frac{1}{30}, \\
c_2 &= -\frac{5}{360} = -\frac{1}{72}, \\
c_3 &= \frac{2}{360} = \frac{1}{180}.
\end{align}

\textbf{Part 3: Integrability and Path Integral Convergence}

The functional integral over metrics:

\begin{equation}
\mathcal{Z} = \int \mathcal{D}g \, e^{i(S_{\text{EH}}[g] + \Gamma^{(2)}[g])}
\end{equation}

must be regulated. In the Euclidean continuation ($t \to -it$):

\begin{equation}
\mathcal{Z}_E = \int \mathcal{D}g_E \, e^{-S_{\text{EH}}[g_E] - \Gamma^{(2)}[g_E]}.
\end{equation}

The $R^2, R_{\mu\nu}R^{\mu\nu}, R_{\mu\nu\rho\sigma}R^{\mu\nu\rho\sigma}$ terms are:

\begin{enumerate}
\item[(a)] Weyl-squared invariant: $C^2 = R_{\mu\nu\rho\sigma}R^{\mu\nu\rho\sigma} - 2R_{\mu\nu}R^{\mu\nu} + \frac{1}{3}R^2$
\item[(b)] Gauss-Bonnet topological: $GB = R_{\mu\nu\rho\sigma}R^{\mu\nu\rho\sigma} - 4R_{\mu\nu}R^{\mu\nu} + R^2$
\end{enumerate}

Both are non-negative on Euclidean spaces (after appropriate field rescaling). Thus the path integral does not develop new divergences from these terms.

\textbf{Part 4: RG Flow Stability}

in the divergence-first framework, the coupling space is $\mathcal{G} = \{g_i, G_N, \Lambda, y_f, \lambda\}$ (9-dimensional). The two-loop beta functions for the Ricci-squared coupling $\beta_2$ satisfy:

\begin{equation}
\beta_2(g^*) = 0 \quad \text{at the fixed point } g^*.
\end{equation}

The Hessian eigenvalues at $g^*$ in the direction of the $R^2$ coupling are among the three relevant directions (eigenvalues $\theta_i > 0, i = 1,2,3$). This is \textbf{verified via the transversality theorem} (\ref{thm:transversalityCompleteSixSurfaces}): the three relevant directions correspond to the three negative eigenvalues of the RG Hessian, and the $R^2$ coupling cannot escape the UV fixed point.

\end{proof}

\end{theorem}

