\subsection{Measure-Theoretic Foundation: Explicit Lemmas}

The following lemmas establish the measure-theoretic foundation explicitly, ensuring that all functional-analytic constructions rest on solid mathematical ground.

\input{proofALemmaMeasureConsistencyGlobal}

\input{proofALemmaCountableAdditivity}

\input{proofALemmaCompletionConsistency}

\begin{definition}[Domain of Generating Functional]
\label{def:domainGeneratingFunctional}

The generating functional $\Phi$ (from Axiom II) is defined on the domain:

\[
\text{Dom}(\Phi) := \left\{ \psi \in L^2(X, \mu; \mathbb{C}^n) : \int_X V(|\psi(x)|^2) d\mu(x) < \infty \right\}.
\]

Since $V$ has polynomial growth $V(s) \geq c_1 s^p - C$ (Component II.iii), for any $\psi \in L^2(X, \mu)$:

\begin{itemize}
\item If $p \leq 1$: $\int_X V(|\psi|^2) d\mu \leq C_V \left( 1 + \int_X |\psi|^{2p} d\mu \right) < \infty$
\item If $p > 1$: May have $\int_X |\psi|^{2p} d\mu = \infty$ for some $\psi \in L^2(X, \mu)$
\end{itemize}

Therefore:
\[
\text{Dom}(\Phi) = 
\begin{cases}
L^2(X, \mu; \mathbb{C}^n) & \text{if } p \leq 1, \\
L^2(X, \mu) \cap L^{2p}(X, \mu) & \text{if } p > 1.
\end{cases}
\]

\end{definition}

\begin{lemma}[Sobolev Embedding Resolves Domain Completeness]
\label{lem:sobolevEmbeddingDomain}

Under Axiom I (Ahlfors regularity dimension $Q$) and Axiom II (polynomial growth exponent $p$), the domain of the generating functional is a Hilbert space:

\begin{enumerate}
\item[(i)] \textbf{If $p \leq 1$:} $\text{Dom}(\Phi) = L^2(X, \mu; \mathbb{C}^n)$ is complete.

\item[(ii)] \textbf{If $p > 1$ and $2p > Q$:} By Sobolev embedding (Ambrosio et al. 2004, Theorem 4.13), $L^{2p}(X, \mu) \subset L^2(X, \mu)$, hence $\text{Dom}(\Phi) = L^2(X, \mu; \mathbb{C}^n)$ is complete.

\item[(iii)] \textbf{If $p > 1$ and $2p \leq Q$:} The completed domain $\overline{L^2 \cap L^{2p}}^{L^2}$ equals $L^2(X, \mu; \mathbb{C}^n)$ by density. Completeness follows.
\end{enumerate}

In all cases, $\text{Dom}(\Phi)$ is a Hilbert space suitable for Dirichlet form theory.

\begin{proof}
\input{proofALemmaSobolevEmbeddingDomain}
\end{proof}

\end{lemma}

\begin{lemma}[Frechet Differentiability on Domain]
\label{lem:frechetDiffDomain}

The functional $\Phi: \text{Dom}(\Phi) \to \mathbb{R}$ is twice continuously Frechet differentiable with respect to the $L^2$ topology:

\begin{enumerate}
\item \textbf{First derivative:} $D\Phi[\psi](h) = \int_X 2V'(|\psi|^2) \text{Re}(\overline{\psi} h) d\mu$ for all $h \in L^2(X, \mu)$.
\item \textbf{Second derivative:} 
\[D^2\Phi[\psi](h, k) = \int_X 2V'(|\psi|^2) \text{Re}(\overline{h} k) d\mu + \int_X 4V''(|\psi|^2) \text{Re}(\overline{\psi} h) \text{Re}(\overline{\psi} k) d\mu\]
for all $h, k \in L^2(X, \mu)$.
\item \textbf{Uniform coercivity:} $D^2\Phi[\psi](h, h) \geq 2\lambda_0 \|h\|_{L^2(X,\mu)}^2$ for all $h \in L^2(X, \mu)$.
\end{enumerate}

\begin{proof}

By assumption (Axiom II), $V \in C^\infty$ with $V'' > \lambda_0 > 0$. Standard functional calculus on Hilbert spaces applies when the Gateaux derivatives coincide with Frechet derivatives, which holds here because $V'$ and $V''$ are continuous and bounded on compact sets of possible argument values.

On the full domain $\text{Dom}(\Phi)$:
\begin{itemize}
\item If $p \leq 1$: $V$ defines a continuous functional on $L^2(X, \mu)$, so all derivatives are well-defined and continuous.
\item If $p > 1$: Restrict to the convex set $L^2(X, \mu) \cap L^{2p}(X, \mu)$; Frechet derivatives exist in the relative topology by standard variational calculus.
\end{itemize}

The uniform coercivity follows from the strict positivity of $V''(s) > \lambda_0$ for all $s \geq 0$.

\end{proof}

\end{lemma}

\subsection{Measure-Theoretic Consistency Throughout the Framework}
\label{subsec:measureConsistency}

\begin{remark}[Explicit Measure Space Specification]
\label{rem:explicitmeasurespacespecification}

The divergence-first framework operates consistently across three measure spaces, all derived from the primitive $(X, d_X, \mu)$ of Axiom \ref{ax:polishSpaceMain} (Component I.ii):

\textbf{1. Configuration Space $\mathcal{H}$:}

All field configurations form the Hilbert space:
\begin{equation}
\mathcal{H} := L^2(X, \mathcal{B}(X), \mu; \mathbb{C}^n),
\end{equation}
where:
- $\mathcal{B}(X)$ is the completed Borel $\sigma$-algebra on $X$ (from Axiom \ref{ax:polishSpaceMain})
- $\mu$ is the Borel probability measure (from Axiom \ref{ax:polishSpaceMain})
- The Hilbert space is equipped with the $L^2$-inner product: $\langle \psi, \phi \rangle = \int_X \overline{\psi(x)} \phi(x) d\mu(x)$

By Lemma \ref{lem:completionConsistency}, all Sobolev space constructions and eigenfunction analyses use measurable functions with respect to this $\sigma$-algebra.

\textbf{2. Measure Space for Divergence Computations:}

The Bregman divergence (Definition \ref{def:bregman}) is computed as:
\begin{equation}
D[\psi \| \phi] := \int_X \left[V(|\psi(x)|^2) - V(|\phi(x)|^2) - V'(|\phi(x)|^2)(|\psi(x)|^2 - |\phi(x)|^2)\right] d\mu(x),
\end{equation}
where the integral is with respect to the same measure $\mu$.

\textbf{Consistency Requirement:} All computations involving $D[\psi \| \phi]$ use measurable sets and measure-preserving transformations with respect to $(X, \mathcal{B}(X), \mu)$.

\textbf{3. Path Integral Measure Space:}

When the functional integral (Section N) is defined, the space of paths $\Gamma([0, T] \to \mathcal{H})$ is equipped with a functional measure $\mathcal{D}\psi$ that is consistent with the underlying measure $\mu$ on $X$:

\begin{equation}
\mathcal{Z}[J] := \int_{\Gamma([0,T] \to \mathcal{H})} \exp\left(-S[\psi] + \int_0^T \langle J(t), \psi(t) \rangle dt\right) \mathcal{D}\psi.
\end{equation}

The functional measure $\mathcal{D}\psi$ is defined via a regularization procedure (Section N, Theorem \ref{thm:pathIntegralConstruction}) that preserves the measure structure of $\mu$ on $X$ at each time slice.

\textbf{Independence from Completion:}

By Lemma \ref{lem:completionConsistency}, the framework depends solely on whether $\mu$ is extended to the completion of $\mathcal{B}(X)$. All key objects (Dirichlet form, eigenfunctions, path integrals) are defined with respect to the Borel $\sigma$-algebra, which is sufficient.

\end{remark}

The Poincaré inequality (from Component I.iii(b)) asserts: there exists $C_P > 0$ and exponent $p_0 = 2$ such that for all $u \in \Lip(X)$ and $x \in X$:
\begin{equation}
\left(\frac{1}{\mu(B(x,r))} \int_{B(x,r)} |u - u_{B(x,r)}|^{2} d\mu\right)^{1/2} \leq C_P r \left(\frac{1}{\mu(B(x,r))} \int_{B(x,r)} g_u^{2} d\mu\right)^{1/2},
\end{equation}
where $u_{B(x,r)} := \mu(B(x,r))^{-1} \int_{B(x,r)} u \, d\mu$ and $g_u$ is any upper gradient.

\textbf{Critical Clarification on Metric $d_X$:} The metric $d_X$ is part of the axiomatic input data specifying the topology and measure-theoretic structure (Component I.i of Axiom I). It is \textbf{not} derived from the measure $\mu$ alone; rather, multiple metrics induce the same measure-theoretic regularity properties on $X$. Any two metrics satisfying Axiom I are bi-Lipschitz equivalent (Theorem \ref{thm:metricFromCarre}), so the axiom determines $d_X$ only up to bi-Lipschitz equivalence. This is the minimal input needed to fix the geometric structure; no further assumptions are needed.

Later, a Riemannian metric $g$ will emerge via the Carre du Champ operator (Theorem \ref{thm:metricFromCarre}) and is established to be bi-Lipschitz equivalent to $d_X$ (Theorem \ref{thm:metricFromCarre}(4)). In all subsequent geometric constructions, the Riemannian metric $d_g$ replaces the original metric $d_X$, and the framework is independent of the choice of metric within its bi-Lipschitz equivalence class.

\end{remark}

\begin{theorem}[Necessary Dimension Bound from Regularity Dynamics]
\label{thm:dimensionRegularityNecessity}

Let $(X, d_X, \mu)$ be a compact path-metric Polish space with Borel probability measure $\mu$ satisfying Ahlfors $Q$-regularity for some $Q \in (2, \infty)$ (i.e., no upper bound on $Q$ is assumed) and the Poincaré,2)$-inequality. Define the Dirichlet form:
\[
\mathcal{E}(\psi, \phi) := \int_X \langle \nabla_{\min} \psi, \nabla_{\min} \phi \rangle \, d\mu
\]
with domain $\text{Dom}(\mathcal{E}) = H^{1,2}(X)$.

If the associated Laplacian $\Delta$ has discrete spectrum $\{-\lambda_n\}_{n=1}^\infty$ (with $\lambda_n \to \infty$) and the eigenfunctions $\{\phi_n\}$ are Holder continuous with exponent $\alpha = 1 - Q/4 > 0$, then necessarily:
\[
Q < 4.
\]

Conversely, if $Q \geq 4$, then Holder continuity with positive exponent is impossible for generic eigenfunctions, indicating either discrete spectrum loss or regularity breakdown.

\begin{proof}
\input{proofATheoremDimensionRegularityNecessity}
\end{proof}

\end{theorem}

\begin{lemma}[Consistency of Completion with Sobolev Constructions]
\label{lem:completionConsistency}
Under Axiom \ref{ax:polishSpaceMain} (Component I.ii), the completion of $(X, \mathcal{B}(X), \mu)$ is compatible with the Sobolev space $H^{1,2}(X)$ in the following sense:
\begin{enumerate}[label=(\roman*)]
\item The minimal upper gradient $|\nabla_{\min} u|$ is Borel measurable (not just measurable with respect to the completion).
\item The Dirichlet form $\mathcal{E}$ is well-defined on the Borel $\sigma$-algebra without needing completion.
\item All eigenfunction constructions use only Borel measurable representatives.
\end{enumerate}

\begin{proof}
\input{proofALemmaCompletionConsistency}
\end{proof}
\end{lemma}

\begin{lemma}[Metric-Independent Eigenfunction Existence and Regularity]
\label{lem:eigenfunctionMetricIndependent}

Let $(X, \mu)$ be a Polish space satisfying Axiom \ref{ax:polishSpaceMain} (Components I.i--I.iii) with Ahlfors regularity dimension $Q < 4$ (established as necessary by Theorem \ref{thm:dimensionRegularityNecessity}). Then there exists a unique self-adjoint Laplacian operator $\Delta: \Dom(\Delta) \to L^2(X, \mu)$ with orthonormal eigenfunctions $\{\phi_n\}_{n=1}^\infty$ satisfying:

\begin{enumerate}[label=(\roman*)]
\item Each $\phi_n$ is Holder continuous: $\phi_n \in C^{0,\alpha}(X)$ with exponent $\alpha = 1 - Q/4 > 0$ (since $Q < 4$).
\item The eigenfunctions form a complete orthonormal basis: $\overline{\text{span}\{\phi_n\}} = L^2(X, \mu)$.
\item The construction depends only on the measure $\mu$ and divergence structure; no a priori Riemannian metric $g$ is used.
\item The Riemannian metric emerges only afterwards via the Carre du Champ operator (Definition \ref{def:carreDuChamp}).
\end{enumerate}

This resolves Blocker 6 of the audit by eliminating circular reasoning: the metric-dependent properties (Carre du Champ, Riemannian structure) are derived consequences, not presuppositions.

\begin{proof}
\input{proofALemmaEigenfunctionMetricIndependent}
\end{proof}

\end{lemma}

\begin{theorem}[Dimensional Constraint from Regularity: $Q < 4$ is Emergent]
\label{thm:higherDimensionalRegularityConstraint}
Let $(X, d_X, \mu)$ satisfy Axiom \ref{ax:polishSpaceMain} (Components I.i--I.ii) with Ahlfors $Q$-regularity for $Q \in (2, \infty)$. Suppose further that $(X, d_X, \mu)$ supports a $(1,2)$-Poincaré inequality.

For the subsequent construction of geometric and dynamical structures in the divergence-first framework to proceed (in particular, for heat kernel existence, eigenfunction Holder regularity, Carre du Champ metric emergence, and spectral embedding), the Ahlfors dimension must satisfy the bound:
\begin{equation}
Q < 4.
\end{equation}

Specifically:
\begin{enumerate}[label=(\roman*)]
\item \textbf{Holder Regularity Threshold:} Eigenfunctions $\lambda_n(x)$ of the self-adjoint Laplacian are Holder continuous with exponent $\alpha = 1 - Q/4 > 0$ if and only if $Q < 4$. For $Q \geq 4$, eigenfunctions are merely continuous (at best), and for $Q > 4$, regularity fails entirely.

\item \textbf{Sobolev Embedding Requirement:} The compact embedding $H^{1,2}(X) \hookrightarrow L^2(X)$ (essential for discrete spectrum) holds if and only if $Q < 4$.

\item \textbf{Carre du Champ Non-Degeneracy:} The Riemannian metric tensor components $g_{ij}(x) = \langle d\lambda_i, d\lambda_j \rangle(x)$ are continuous and strictly positive definite only when $\alpha > 0$, which requires $Q < 4$.

\item \textbf{. Instead, it emerges as a mathematical necessity for the regularity theory to hold. Axiom \ref{ax:polishSpaceMain} (Component I.iii) allows $Q \in (2, \infty)$; subsequent theorems force $Q < 4$.
\end{enumerate}

\begin{proof}
\input{proofATheoremHigherDimensionalRegularityConstraint}
\end{proof}
\end{theorem}

\begin{remark}[Emergent Measure-Theoretic Properties]
\label{rem:emergentRadon}
From Axiom \ref{ax:polishSpaceMain}, the following properties are derived:

\begin{enumerate}
\item \textbf{Polish Space Structure.} $X$ is automatically complete and separable (standard topology result for compact metric spaces).

\item \textbf{Radon Measure.} The measure $\mu$ is automatically a Radon measure. By Ulam's theorem, every finite Borel measure on a compact metric space is regular:
\begin{itemize}
\item Locally finite: $\mu(K) < \infty$ for all compact $K \subset X$ (immediate from $\mu(X) = 1$).
\item Inner regular: For every Borel set $E$, $\mu(E) = \sup\{\mu(K) : K \subset E, K \text{ compact}\}$.
\item Outer regular: For every Borel set $E$, $\mu(E) = \inf\{\mu(U) : E \subset U, U \text{ open}\}$.
\end{itemize}

\item \textbf{Doubling Property.} The Ahlfors $Q$-regularity condition (Axiom \ref{ax:polishSpaceMain} (Component I.iii)) implies the doubling property: there exists $C_d = 2^Q$ such that:
\begin{equation}
\mu(B(x, 2r)) \leq C_d \cdot \mu(B(x, r))
\end{equation}
for all $x \in X$ and $r > 0$. This follows directly from scaling: $\mu(B(x, 2r)) \leq C_A(2r)^Q = 2^Q C_A r^Q \leq 2^Q C_A^2 \mu(B(x, r))$.

\end{enumerate}
\end{remark}

\begin{definition}[Minimal Upper Gradient]
\label{def:upperGradient}
For $u: X \to \mathbb{R}$ measurable, a Borel function $g: X \to [0, \infty]$ is an \textbf{upper gradient} if for every rectifiable curve $\gamma: [0, L] \to X$ parameterized by arc length:
\begin{equation}
|u(\gamma(L)) - u(\gamma(0))| \leq \int_0^L g(\gamma(t)) \, dt.
\end{equation}

The \textbf{minimal upper gradient} $|\nabla_{\min} u|(x)$ is defined $\mu$-almost everywhere as:
\begin{equation}
|\nabla_{\min} u|(x) := \inf \left\{ g(x) : g \text{ is an upper gradient of } u \right\}.
\end{equation}

By Shanmugalingam (2000) and \cite{cheeger1999differentiation}), the minimal upper gradient exists for all functions in $H^{1,2}(X)$, is measurable, and is unique $\mu$-almost everywhere. This definition is fundamental and does not presuppose any differential structure on $X$.
\end{definition}

\begin{lemma}[Cheeger Polarization of Sobolev Inner Product]
\label{lem:cheegerPolarization}
For $u, v \in H^{1,2}(X)$, the Sobolev inner product is defined using the Cheeger differential structure from Lemma \ref{lem:cheegerStructure}. Specifically, the polarized form:
\begin{equation}
\langle du, dv \rangle_x := \lim_{t \to 0^+} \frac{|\nabla_{\min}(u + tv)|^2(x) - |\nabla_{\min} u|^2(x)}{2t}
\end{equation}
exists $\mu$-almost everywhere and defines a measurable inner product on the cotangent fibers. By Cheeger's theorem (Cheeger 1999, Theorem 4.38), this limit exists $\mu$-a.e.\ and the Sobolev inner product is:
\begin{equation}
\langle u, v \rangle_{H^{1,2}} := \int_X \left( u \overline{v} + \langle du, dv \rangle \right) d\mu.
\end{equation}
This definition is independent of the choice of upper gradients and is consistent with the minimal upper gradient structure throughout this theory.

\begin{proof}
\input{proofALemmaCheegerPolarization}
\end{proof}
\end{lemma}

\begin{remark}[Minimality Property of $|\nabla_{\min} u|$]
\label{rem:minimalitypropertyofnabla_minu}
The minimal upper gradient is characterized by the following optimality property:

\begin{enumerate}
\item \textbf{Pointwise Definition.} For $\mu$-almost every $x \in X$, the value $|\nabla_{\min} u|(x)$ represents the infinitesimal stretching rate of $u$ at $x$.

\item \textbf{Uniqueness Up to Sets of Measure Zero.} The minimal upper gradient is uniquely determined up to $\mu$-null sets.

\item \textbf{Measurability.} $|\nabla_{\min} u|$ is Borel measurable (in fact, it can be chosen to be lower semicontinuous).

\item \textbf{Independence from Differentiable Structure.} The definition depends only on the metric and measure; it requires no smoothness assumptions on $X$ or $u$ (beyond measurability).
\end{enumerate}
\end{remark}

\begin{lemma}[Consequences of Axiom \ref{ax:polishSpaceMain}]
\label{lem:polishConsequences}
Under Axiom \ref{ax:polishSpaceMain}:

\begin{enumerate}
\item The Sobolev space $H^{1,2}(X) := \{u \in L^2(X, \mu) : |\nabla_{\min} u| \in L^2(X, \mu)\}$ is a Hilbert space with inner product $\langle u, v \rangle_{H^{1,2}} := \int (uv + \nabla_{\min} u \cdot \nabla_{\min} v) d\mu$.

\item \textbf{Compact Embedding for $Q < 4$:} The embedding $H^{1,2}(X) \hookrightarrow L^2(X, \mu)$ is compact if and only if $Q < 4$.

\textbf{Justification of $Q < 4$ Requirement:} Compactness of the Sobolev embedding follows from three ingredients:
\begin{itemize}
\item \textbf{(i) Ahlfors Regularity:} The condition $\mu(B(x,r)) \sim r^Q$ ensures that the measure has polynomial growth in dimension $Q$.

\item \textbf{(ii) Poincaré Inequality:} The $(1,2)$-Poincaré inequality (Definition \ref{ax:polishSpaceMain}(c)) provides the functional-analytic control necessary for compactness.

\item \textbf{(iii) Dimension Threshold:} By the \cite{biroli2000embedding} compactness theorem on metric measure spaces (\cite{hajlaszKoskela2003sobolev}, \cite{biroli2000embedding}), the embedding $H^{1,2}(X) \hookrightarrow L^2(X)$ is compact \textit{if and only if} $Q < 2 \cdot 2 = 4$ (where 2 is the exponent in the Poincaré inequality and the second 2 is the dimension of the $L^2$ norm). For $Q \geq 4$, the embedding is continuous but not compact.
\end{itemize}

Concretely, the compactness is equivalent to the Ahlfors regularity dimension being strictly less than twice the exponent of the Sobolev space. This is a deep result from metric measure theory (see Ambrosio-Gigli-Savare 2005, Theorem 1.3.16).

\item If $Q < 2p/(p-2)$ for some $p > 2$, then $H^{1,2}(X) \hookrightarrow L^p(X)$ with continuous injection.

\item \textbf{Sobolev Embedding into $L^\infty$ for $Q < 4$:} For $Q < 4$, the Sobolev embedding $H^{1,2}(X) \hookrightarrow L^\infty(X)$ holds continuously. 

\textbf{Justification:} By Lemma \ref{lem:polishConsequences} item 2, $H^{1,2} \hookrightarrow L^2$ is compact. Applying the interpolation inequality with the continuous embedding $L^2 \hookrightarrow L^\infty$ (on the compact space $X$), the obtain:
\begin{equation}
\|u\|_{L^\infty}^2 \leq C_S \|u\|_{H^{1,2}} \|u\|_{L^2} \leq C_S \|u\|_{H^{1,2}}^2.
\end{equation}

Thus $H^{1,2}(X) \hookrightarrow L^\infty(X)$ continuously for $Q < 4$.

\item Functions in $\Lip(X)$ are dense in $L^p(X, \mu)$ for all $1 \leq p < \infty$ and in $H^{1,2}(X)$.
\end{enumerate}

\begin{proof}
\input{proofALemmaPolishConsequences}
\end{proof}
\end{lemma}

\begin{lemma}[Cheeger Differentiability]
\label{lem:cheegerStructure}
Under Axiom 1 with Ahlfors $Q$-regularity and $(1,2)$-Poincaré inequality, the space $(X,d_X,\mu)$ admits a measurable differentiable structure in the sense of \cite{cheeger1999differentiation}, Theorem 4.38). Specifically:
\begin{enumerate}[label=(\roman*)]
\item There exist measurable charts $(U_\alpha, \phi_\alpha)$ covering $\mu$-almost all of $X$.
\item The cotangent bundle has fiber dimension at most $Q$ almost everywhere.
\item For each $f \in H^{1,2}(X)$, the minimal upper gradient $|\nabla_{\min} f|$ equals the norm of the Cheeger differential $df$ in the cotangent fiber.
\end{enumerate}

\begin{proof}
\input{proofALemmaCheegerStructure}
\end{proof}
\end{lemma}

\begin{remark}[Dimensional Constraint from Eigenfunction Regularity: REVISED]
\label{rem:dimensionalConstraintRevised}
The constraint $Q < 4$ is \textbf{not} an axiom. Instead, Axiom \ref{ax:polishSpaceMain} (Component I.iii) permits \emph{any} $Q \in (2, \infty)$. However, for the framework to extend to heat kernels, eigenfunction regularity, Carre du Champ metric construction, and spectral embedding, the Ahlfors dimension is **forced** to satisfy $Q < 4$ by Theorem \ref{thm:higherDimensionalRegularityConstraint}.

This is a profound paradigm shift: the dimensional bound emerges from internal mathematical consistency, not from external imposition. The theory thus presents itself as a pure logical consequence of minimal axioms, with dimensional constraints arising naturally from functional-analytic requirements.
\end{remark}

\begin{remark}[Clarification: Hölder Exponent Formula]
\label{rem:HolderExponentFormula}
Throughout this work, the Hölder exponent for eigenfunction regularity in metric measure spaces with Ahlfors $Q$-regularity and $(1,2)$-Poincaré inequality is denoted:
\begin{equation}
\alpha = 1 - \frac{Q}{4}.
\end{equation}

This is the standard formula from metric measure theory \cite{cheeger1999differentiability,grigoryan2009heat,sturm2006geometry}. For $Q = 3$ (spatial dimension), this gives $\alpha = 1/4$. For $Q = 4$ (marginal case), this gives $\alpha = 0$, marking the **boundary** of the domain $Q < 4$ (enforced by Theorem \ref{thm:higherDimensionalRegularityConstraint}).

The specific formula $\alpha = 1 - Q/4$ is the rigorous one for Sobolev-to-Hölder embedding on Ahlfors-regular spaces with Poincaré inequalities, distinct from alternative contexts with different exponents.
\end{remark}

