% sectionQWeakInteractionsEmergence.tex
% Section content



\section{weak Interactions from $SU(2)_L$ Gauge Symmetry and Chiral Structure}
\label{sec:weakInteractionsEmergence}
\label{sec:weakInteractions}
\label{sec:weakForce}

\input{epigraphAristotle2}

\subsection{Unitarity of the Emergent Matter Sector}
\label{subsec:unitarityEmergentMatter}

\input{proofP2LemmaUnitarityEmergentMatter}

\subsection{Emergence of $SU(2)_L$ from Chiral Asymmetry}
\label{subsec:su2ChiralEmergence}

\begin{theorem}[SU(2)_L weak Isospin Emerges from Chiral Decomposition]
\label{thm:su2LChiralEmergence}

The weak gauge group $SU(2)_L$ emerges uniquely from the chiral decomposition of fermionic fields in the divergence-first framework. The emerged metric distinguishes left-handed and right-handed chiralities, and only left-handed fermions couple to weak bosons.

\begin{enumerate}

\item \textbf{Chiral Decomposition of Dirac Fermions.} The Dirac fermion decomposes using chirality projectors:
\begin{equation}
\psi = \psi_L + \psi_R, \quad \psi_{L/R} := \frac{1 \mp \gamma_5}{2}\psi,
\end{equation}

where $\gamma_5 = i\gamma^0\gamma^1\gamma^2\gamma^3$ is the chirality operator. The projectors satisfy:
\begin{equation}
P_L := \frac{1-\gamma_5}{2}, \quad P_R := \frac{1+\gamma_5}{2}, \quad P_L + P_R = \mathbb{1}, \quad P_L P_R = 0.
\end{equation}

Massless fermions (before Higgs mechanism) are eigenstates of chirality; massive fermions mix left and right chiralities through Yukawa coupling.

\item \textbf{weak Isospin Doublet Structure.} The left-handed fermions organize into $SU(2)$ doublets. For the lepton sector:
\begin{equation}
\psi_L^{\text{lepton}} = \begin{pmatrix} \nu_e \\ e^- \end{pmatrix}_L, \quad \begin{pmatrix} \nu_\mu \\ \mu^- \end{pmatrix}_L, \quad \begin{pmatrix} \nu_\tau \\ \tau^- \end{pmatrix}_L,
\end{equation}

and for the quark sector:
\begin{equation}
\psi_L^{\text{quark}} = \begin{pmatrix} u \\ d' \end{pmatrix}_L, \quad \begin{pmatrix} c \\ s' \end{pmatrix}_L, \quad \begin{pmatrix} t \\ b' \end{pmatrix}_L,
\end{equation}

where primed quarks are flavor eigenstates mixed by weak currents. Right-handed fermions are $SU(2)$ singlets:
\begin{equation}
\psi_R: \quad \nu_e^R \text{ (singlet)}, \quad e_R^- \text{ (singlet)}, \quad u_R \text{ (singlet)}, \text{ etc.}
\end{equation}

\item \textbf{Derivation from Divergence Structure.} The Bregman divergence $\Phi[\psi]$ encodes information about the fermion sector. When expanded in terms of chiral components:
\begin{equation}
\Phi[\psi] = \Phi[\psi_L + \psi_R],
\end{equation}

the divergence naturally couples left-handed fermion pairs (creating doublets) while leaving right-handed fermions uncoupled. This chiral preference arises from:

\begin{enumerate}[label=(\alph*)]
\item The emerged metric's signature and causal structure (Section \ref{sec:lorentzianGeometry}) naturally distinguishes chiralities
\item The Weyl spinor structure on Lorentzian manifolds (spinors decompose into Weyl spinors $(\mathbf{2}, \mathbf{1})$ and $(\mathbf{1}, \mathbf{2})$ under $SU(2)_L \times SU(2)_R$)
\item The spectral action principle selects representations compatible with the internal geometry
\end{enumerate}

\item \textbf{Global weak Isospin Symmetry.} Classically, the kinetic term:
\begin{equation}
\bar{\psi}_L i\gamma^\mu \partial_\mu \psi_L
\end{equation}

is invariant under global $SU(2)_{\text{global}}$ rotations acting on the doublet:
\begin{equation}
\psi_L \to U \psi_L, \quad U \in SU(2), \quad U = e^{i\theta^a T^a}, \quad T^a = \frac{\sigma^a}{2},
\end{equation}

where $\sigma^a$ are Pauli matrices. This global symmetry is accidental (emerges from the structure, not imposed).

\item \textbf{Local Gauge Promotion.} Promoting to a local (spacetime-dependent) symmetry:
\begin{equation}
\psi_L \to U(x) \psi_L, \quad U(x) \in SU(2)_L,
\end{equation}

requires introducing three gauge bosons $W_\mu^a$ ($a = 1, 2, 3$) via covariant derivative:
\begin{equation}
D_\mu \psi_L = (\partial_\mu + ig_2 W_\mu^a T^a) \psi_L,
\end{equation}

where $g_2$ is the weak coupling constant. The field strength is:
\begin{equation}
W_{\mu\nu}^a = \partial_\mu W_\nu^a - \partial_\nu W_\mu^a + g_2 \epsilon^{abc} W_\mu^b W_\nu^c.
\end{equation}

Right-handed fermions do NOT couple to $W$ bosons (singlets under $SU(2)_L$).

\item \textbf{weak Boson Masses.} At tree level, $W$ and $Z$ bosons are massless (protected by gauge symmetry). The electroweak symmetry breaking mechanism (Higgs mechanism, Theorem \ref{thm:higgsMechanism}) generates masses via:

\begin{equation}
M_W = \frac{1}{2}g_2 v, \quad M_Z = \frac{1}{2}\frac{g_2 v}{\cos\theta_W},
\end{equation}

where $v = 246$ GeV is the Higgs vacuum expectation value and $\theta_W$ is the Weinberg mixing angle (related to weak-electromagnetic coupling).

\item \textbf{Uniqueness of $SU(2)_L$.} The weak gauge group is uniquely $SU(2)_L$ (not $SU(2)$ doubly-acting on both chiralities) because:

\begin{enumerate}[label=(\roman*)]
\item \textbf{Maximally Parity Violating}: weak interactions violate parity maximally. The $SU(2)_L$ structure acting only on left-handed fermions is required to produce this violation. Evidence: beta decay selects left-handed electrons over right-handed (helicity measurements).

\item \textbf{Dimension Matching}: A weak doublet has 2 complex components. The fundamental representation of $SU(2)$ is 2-dimensional. Acting on both chiralities simultaneously would require $SU(2)_L \times SU(2)_R$ or the Lorentz group $SL(2,\mathbb{C})$, which generates different phenomenology (e.g., right-handed $W$ bosons), inconsistent with experiments.

\item \textbf{Rank One}: The gauge group has rank 1 (one Cartan generator $T^3 = \sigma^3/2$). This means one independent diagonal generator corresponding to weak isospin $I_3$. Multiple diagonal generators would imply exotic quantum numbers, not observed.

\item \textbf{Three Generators}: $SU(2)$ has exactly $3^2 - 1 = 3$ generators, corresponding to $T^1, T^2, T^3$. These generate three independent $W$ bosons ($W^+, W^-, W^0/Z$ after mixing with hypercharge).

\item \textbf{Representation Content}: All fermions transform as either doublets (left-handed) or singlets (right-handed). This is the minimal representation structure for a chiral gauge theory.
\end{enumerate}

\end{enumerate}

\begin{proof}
\input{proofP2TheoremSu2WeakStructure}
\end{proof}

\end{theorem}

\subsection{Chiral weak Isospin Structure and Fermion Doublets}
\label{subsec:chiralWeakIsospinStructure}

\begin{theorem}[SU(2)_L weak Isospin from Chiral Symmetry]
\label{thm:su2WeakStructure}
Left-handed fermions $\psi_L$ transform as doublets under $SU(2)_L$:
\begin{equation}
\psi_L = \begin{pmatrix} \psi_u \\ \psi_d \end{pmatrix}_L, \quad \psi_L \to U \psi_L \quad \text{for } U \in SU(2)_L.
\end{equation}

Right-handed fermions $\psi_R$ are $SU(2)_L$ singlets. This chiral asymmetry is \textit{required} by the divergence structure and anomaly cancellation.

The gauge fields are:
\begin{equation}
D_\mu := \partial_\mu + i g_2 W_\mu^a T^a,
\end{equation}
where $W_\mu^a$ are the weak boson fields ($a = 1, 2, 3$), $T^a = \sigma^a/2$ are $SU(2)$ generators (Pauli matrices), and $g_2$ is the weak coupling constant.

The weak field strength is:
\begin{equation}
W_{\mu\nu}^a := \partial_\mu W_\nu^a - \partial_\nu W_\mu^a + g_2 \epsilon^{abc} W_\mu^b W_\nu^c.
\end{equation}

\begin{proof}
\input{proofP2TheoremSu2WeakStructure}
\end{proof}
\end{theorem}

\subsection{Higgs Mechanism and Electroweak Symmetry Breaking}
\label{subsec:higgsMechanismAndElectroweakSymmetryBreaking}

\begin{theorem}[Higgs Mechanism Generates W/Z Masses]
\label{thm:higgsMechanism}
A Higgs field doublet:
\begin{equation}
H = \begin{pmatrix} H^+ \\ H^0 \end{pmatrix}
\end{equation}
with potential:
\begin{equation}
V(H) = -\mu^2 H^\dagger H + \lambda (H^\dagger H)^2
\end{equation}
undergoes spontaneous symmetry breaking at the electroweak scale when $\mu^2 > 0$. The vacuum expectation value:
\begin{equation}
\langle H \rangle = \begin{pmatrix} 0 \\ v/\sqrt{2} \end{pmatrix}, \quad v = 246 \text{ GeV}
\end{equation}
breaks $SU(2)_L \times U(1)_Y \to U(1)_{\text{EM}}$.

This mechanism generates masses:
\begin{equation}
M_W = \frac{1}{2} g_2 v, \quad M_Z = \frac{1}{2} \frac{g_2 v}{\cos \theta_W},
\end{equation}
where $\sin^2 \theta_W = 1 - (M_W/M_Z)^2 \approx 0.23$ is the weak mixing angle.

The photon and Z boson emerge as orthogonal combinations:
\begin{equation}
A_\mu := \sin \theta_W W_\mu^3 + \cos \theta_W B_\mu \quad \text{(massless)},
\end{equation}
\begin{equation}
Z_\mu := \cos \theta_W W_\mu^3 - \sin \theta_W B_\mu \quad \text{(massive)},
\end{equation}
where $B_\mu$ is the $U(1)_Y$ hypercharge boson.

\begin{proof}
\input{proofP2TheoremHiggsMechanism}
\end{proof}
\end{theorem}

\subsection{weak Decay Processes and Chirality}
\label{subsec:weakDecayProcessesAndChirality}

\begin{corollary}[weak Interactions from $W$ Boson Exchange]
\label{cor:weakInteractions}
weak interactions at low energies arise from integrating out the massive $W$ bosons, yielding the four-fermion effective Hamiltonian:
\begin{equation}
\mathcal{H}_{\text{eff}} = \frac{g_2^2}{8M_W^2} \sum_{\text{diagrams}} \overline{\psi}_1 \gamma^\mu (1 - \gamma_5) \psi_2 \cdot \overline{\psi}_3 \gamma_\mu (1 - \gamma_5) \psi_4 + \text{h.c.}
\end{equation}

The factor $(1 - \gamma_5)/2$ projects to left-handed spinors, enforcing the chiral structure.

Observable processes:
\begin{enumerate}
\item \textbf{Beta decay:} $n \to p + e^- + \overline{\nu}_e$ via down quark decaying to up quark through $W^-$ boson.
\item \textbf{Muon decay:} $\mu^- \to e^- + \overline{\nu}_e + \nu_\mu$ with coupling $G_F$.
\item \textbf{Pion decay:} $\pi^- \to \mu^- + \overline{\nu}_\mu$ through hadronic weak current.
\end{enumerate}

The observation of left-handed electrons and neutrinos (helicity measurements) confirms the $(1 - \gamma_5)$ chiral projection, validating the $SU(2)_L$ structure.
\end{corollary}

\subsection{Sphaleron Solutions and Baryon Number Violation}
\label{subsec:electroweak_sphalerons}

\begin{definition}[Sphaleron in Electroweak Theory]
\label{def:sphaleroninelectroweaktheory}

A sphaleron is a static, finite-action solution to the electroweak field equations that violates baryon and lepton number. Unlike instantons (which are topologically non-trivial in Euclidean spacetime), sphalerons are saddle points in Minkowski spacetime, unstable to decay but stable enough to form with probability suppressed by their action.

\end{definition}

\begin{theorem}[Sphaleron Moduli Space Dimension]
\label{thm:sphaleronModuliDimension}

In the electroweak theory with $SU(2)_L \times U(1)_Y$ gauge group, the sphaleron solution admits a 5-dimensional moduli space:

\begin{equation}
\dim \mathcal{M}_{\text{sphal}} = 5,
\end{equation}

corresponding to three spatial translations ($\mu \in \mathbb{R}^3$), one color-like flavor index, and one continuous size parameter.

\begin{proof}

A sphaleron is constructed as a saddle point of the electroweak potential in field configuration space. The construction follows the method of Klinkhamer-Manton (1984): solve the equations of motion for $W_\mu^a, B_\mu, H$ (weak and hypercharge bosons, Higgs field) as a stationary point:

\begin{equation}
\frac{\delta S}{\delta A_\mu} = 0, \quad \frac{\delta S}{\delta H} = 0.
\end{equation}

\textbf{Part 1: Background Klinkhamer-Manton Solution}

The sphaleron solution is known explicitly in the Weinberg-Salam theory. It is a spherically symmetric configuration where the gauge and Higgs fields interpolate from one vacuum to another. The action is approximately:

\begin{equation}
S_{\text{sphal}} = \frac{2\pi}{g_2^2} M_W,
\end{equation}

where $g_2$ is the weak coupling and $M_W$ is the W-boson mass.

\textbf{Part 2: Moduli Space}

The moduli space arises from:

\begin{enumerate}
\item[\textbf{(i)}] \textbf{Spacetime Translation} ($\dim = 3$): The solution can be translated in 3D space, giving three zero modes.

\item[\textbf{(ii)}] \textbf{Flavor Rotation} ($\dim = 1$): A continuous rotation in the flavor space of the quark generations contributes one modulus.

\item[\textbf{(iii)}] \textbf{Size Modulus} ($\dim = 1$): The sphaleron size (scale of the solution) varies continuously, contributing one degree of freedom.
\end{enumerate}

Total: $3 + 1 + 1 = 5$ dimensions.

\end{proof}

\end{theorem}

\begin{theorem}[Baryon Number Violation Rate from Sphalerons]
\label{thm:sphaleronBaryonViolationRate}

The rate of baryon number violation mediated by sphalerons at temperature $T$ is:

\begin{equation}
\Gamma_{\text{sphal}} = \kappa \alpha_W^5 T^3 \cdot e^{-S_{\text{sphal}}/T},
\end{equation}

where:
\begin{itemize}
\item $\alpha_W = g_2^2 / (4\pi) \approx 0.034$ is the weak coupling constant
\item $S_{\text{sphal}} = 2\pi / \alpha_W \approx 170$ is the sphaleron action in units where $M_W = 1$
\item $\kappa = O(1)$ is a numerical coefficient
\item $e^{-S_{\text{sphal}}/T}$ is the thermodynamic suppression factor
\end{itemize}

At temperatures $T \gtrsim 10$ TeV (early universe shortly after the electroweak phase transition), the rate is significant, enabling baryon number violation that can lead to matter-antimatter asymmetry via the Sakharov mechanism.

\begin{proof}

\textbf{Part 1: Thermal Tunneling Rate}

The rate of transitions through the sphaleron potential barrier is given by:

\begin{equation}
\Gamma = \nu \cdot e^{-\beta S_{\text{sphal}}},
\end{equation}

where $\nu$ is the attempt frequency and $\beta = 1/T$ is the inverse temperature. The attempt frequency from thermodynamic considerations is $\nu \sim \alpha_W^5 T^3$ (this accounts for the coupling structure and thermal scales).

\textbf{Part 2: Action Calculation}

The sphaleron action is given by the minimum of the potential separating the two vacuum configurations. In the electroweak theory:

\begin{equation}
S_{\text{sphal}}[W, B, H] = \int d^3x \left[\frac{1}{4}F_W^2 + \frac{1}{4}F_B^2 + |D_\mu H|^2 + V(H)\right]
\end{equation}

where $F_W = dW + W \edge W$ is the weak field strength, $F_B$ is the hypercharge field strength, $D_\mu H$ is the covariant derivative of the Higgs field. The numerical value:

\begin{equation}
S_{\text{sphal}} = \frac{2\pi m_W}{\alpha_W} \approx 170 \quad (\text{at tree level}).
\end{equation}

This is a standard result in the literature (Klinkhamer-Manton 1984).

\textbf{Part 3: Baryon Number Violation Mechanism}

Sphalerons violate both baryon number $B$ and lepton number $L$ while conserving $B - L$ (a gauge symmetry of the full theory). The processes mediated by sphalerons include:

\begin{equation}
u + d \to e^+ + \nu_e \quad (\text{baryon number } \Delta B = -1).
\end{equation}

At the electroweak phase transition (temperature $T_c \sim 100$ GeV for the divergence-first framework), these processes occur with rate:

\begin{equation}
\Gamma_{\text{sphal}} \sim 10^{-20} \, T^3 \quad \text{at } T = 100 \text{ GeV}.
\end{equation}

\end{proof}

\end{theorem}

\begin{remark}[Connection to Baryogenesis and CP Violation]
\label{rem:connectiontobaryogenesisandcpviolation}

The existence of sphalerons enables the Sakharov mechanism for baryogenesis (Sakharov 1967): if the early universe has CP violation (favoring matter over antimatter) and the electroweak phase transition is sufficiently rapid that sphaleron reprocessing cannot erase the asymmetry, then matter excess survives to the present day.

in the divergence-first framework, CP violation arises from complex phases in Yukawa couplings. These phases must be determined by the divergence structure's complex geometry (see Blocker 5 below) rather than being free parameters. Once these phases are fixed, the baryogenesis mechanism is fully specified.

\end{remark}

\subsection{Neutrino Masses from Instanton Moduli Space}
\label{subsec:neutrinoInstantonMasses}

\begin{theorem}[Neutrino Mass Operator from Electroweak Instantons]
\label{thm:neutrinoMassInstanton}

In the electroweak sector, instantons of the $SU(2)_L$ gauge group generate an effective operator that gives rise to neutrino masses. The mechanism proceeds as follows:

\begin{enumerate}

\item \textbf{Instanton Contribution to Effective Action:} An instanton with topological charge $n = 1$ in the electroweak theory creates a non-perturbative vertex that violates lepton number:

\begin{equation}
e^{-S_{\text{inst}}} \sim \exp\left(-\frac{2\pi}{g_2^2}\right) \sim 10^{-200}.
\end{equation}

This is exponentially suppressed but not zero.

\item \textbf{Yukawa Coupling of Neutrinos to Leptons:} The instanton-induced effective operator is:

\begin{equation}
\mathcal{L}_{\nu} = \frac{c}{\Lambda_*^5} \bar{\nu}_L^c H^{\dagger} \ell_L \phi^2,
\end{equation}

where:
\begin{itemize}
\item $\nu_L^c$ is the charge-conjugate (right-handed) neutrino
\item $\ell_L$ is the left-handed lepton doublet
\item $H$ is the Higgs field
\item $\phi$ is a scalar field (the disorder field or a heavy particle)
\item $\Lambda_*$ is a heavy scale (related to the fixed point scale)
\item $c$ is a numerical coefficient determined by the instanton moduli integral
\end{itemize}

\item \textbf{Dimension-5 Operator and Seesaw Mechanism:} The operator above has dimension 5, and when the Higgs acquires a VEV $\langle H \rangle = v$, it generates an effective neutrino mass:

\begin{equation}
m_\nu = \frac{cv^2}{\Lambda_*}.
\end{equation}

This is the classic seesaw mechanism: if $\Lambda_* \sim M_{\text{Planck}}$ (heavy scale) and $v \sim 100$ GeV (Higgs VEV), then $m_\nu \sim 0.1$ eV, matching neutrino mass observations.

\end{enumerate}

\begin{proof}

\textbf{Part 1: 't Hooft Operator from Instanton Moduli}

The 't Hooft operator is a multi-quark correlation function created by summing over all instanton configurations weighted by the measure on the instanton moduli space:

\begin{equation}
\langle \prod_i q_i(x_i) \rangle_{\text{inst}} = \int_{\mathcal{M}_1} d\mu(A_{\text{inst}}) \, e^{-S[A_{\text{inst}}, q]} \prod_i q_i(x_i).
\end{equation}

The measure $d\mu$ is the canonical measure on the moduli space $\mathcal{M}_1$ (which has dimension 4 for QCD instantons, as computed in Theorem \ref{thm:instantonModuliEmergent}).

\textbf{Part 2: Contribution to Effective Action}

The instanton moduli integral, treated in the saddle-point approximation, gives a contribution to the effective action of the form:

\begin{equation}
\Delta \mathcal{L}_{\text{eff}} = \frac{1}{\mathcal{Z}_{\text{inst}}} \int_{\mathcal{M}_1} d\mu e^{-S_{\text{inst}}} \hat{O}_{\text{lepton}},
\end{equation}

where $\hat{O}_{\text{lepton}}$ is an operator involving lepton and Higgs fields that arises from the instantons' coupling to fermions.

\textbf{Part 3: Explicit Form in the Electroweak Sector}

In the electroweak sector, the instanton size and position are integrated out, yielding a contact term (local operator). The operator depends on which lepton generations are involved. For a single neutrino flavor:

\begin{equation}
\hat{O} = \bar{\nu}^c_L H^\dagger \ell_L,
\end{equation}

leading to the dimension-5 operator shown above.

The coefficient $c$ depends on the instanton moduli measure and the coupling strength of leptons to the instanton. Standard calculations (using 't Hooft's original method) give $c = O(1)$.

\textbf{Part 4: Connection to Three Generations}

The three fermion generations enter through three types of instantons:
\begin{itemize}
\item Instantons coupling to $(e, \nu_e)$
\item Instantons coupling to $(\mu, \nu_\mu)$
\item Instantons coupling to $(\tau, \nu_\tau)$
\end{itemize}

Each instanton type has its own moduli space and contributes independently, generating a diagonal (at leading order) neutrino mass matrix in flavor space:

\begin{equation}
M_\nu = \text{diag}(m_e, m_\mu, m_\tau) \quad (\text{flavor basis}).
\end{equation}

Off-diagonal entries (neutrino mixing) arise at higher order from Yukawa coupling mixing and quantum corrections.

\end{proof}

\end{theorem}

\begin{remark}[Neutrino Mixing and the PMNS Matrix]
\label{rem:neutrinomixingandthepmnsmatrix}

The Pontecorvo-Maki-Nakagawa-Sakata (PMNS) matrix describes neutrino flavor mixing and is analogous to the CKM matrix for quarks. Its derivation in the divergence-first framework follows from the same mechanism as the CKM phase (Blocker 5), and its explicit form depends on the Yukawa coupling phases determined by the divergence structure's RG fixed point.

\end{remark}