% Part of sectionCDirichletFormTheory.tex
\subsection{Kinetic Energy as Intrinsic Geometric Property of Polish Spaces}
\label{subsec:kineticEnergyIntrinsic}

\begin{theorem}[Kinetic Energy Structure from Metric Measure Space - Intrinsic to Axiom I]
\label{thm:kineticEnergyIntrinsicToMetric}

The kinetic energy structure of the Dirichlet form is \textbf{not} a hidden axiom. It emerges as a canonical geometric invariant from the Polish space metric-measure structure (Axiom \ref{ax:polishSpace}) alone.

\begin{enumerate}[label=(\roman*)]

\item \textbf{Minimal Upper Gradient is Intrinsic:} On any metric measure space $(X, d, \mu)$ satisfying Ahlfors $Q$-regularity and Poincaré inequality (Axiom I), the minimal upper gradient $\nabla_{\min}$ is uniquely defined without external axiomatization. It depends only on the metric structure $(X, d, \mu)$.

\item \textbf{Kinetic Energy Form:} The quadratic form:
\begin{equation}
\mathcal{E}_{\text{kin}}[\psi] := \int_X |\nabla_{\min} \psi|^2 \, d\mu
\end{equation}
is a \textbf{canonical geometric invariant} arising directly from the metric structure. This form does \textbf{NOT} require external axiomatization (no ``Axiom III'').

\item \textbf{Foundation in Metric Theory:} By Cheeger's 1999 metric Sobolev space theory and subsequent developments (Hajłasz, Koskela, Shanmugalingam), the Sobolev space $H^{1,2}(X)$ is well-defined on metric measure spaces via:
\begin{equation}
\|f\|_{H^{1,2}}^2 := \|f\|_{L^2}^2 + \int_X |\nabla_{\min} f|^2 \, d\mu.
\end{equation}
This construction is intrinsic to metric geometry and requires no external structure.

\item \textbf{Coercivity for $Q < 4$:} For Ahlfors $Q$-regular spaces with Poincaré inequality, the kinetic energy form is coercive when $Q < 4$. Specifically, there exists $\lambda_0 > 0$ such that:
\begin{equation}
\mathcal{E}_{\text{kin}}[\psi] \geq \lambda_0 \|\psi\|_{H^{1,2}}^2.
\end{equation}
This bound depends only on the Ahlfors dimension $Q$ and the Poincaré constant (both intrinsic to the Polish space).

\item \textbf{Logical Acyclicity:} The emergence of kinetic structure is acyclic:
\begin{center}
Polish space $(X, d, \mu)$ [Axiom I]
$\to$ Minimal upper gradient $\nabla_{\min}$
$\to$ Kinetic form $\mathcal{E}_{\text{kin}}[\psi]$
$\to$ Sobolev space $H^{1,2}(X)$
$\to$ [No circularity]
\end{center}

The Bregman divergence quadratic form (Axiom II) is added \textit{independently} to yield the complete Dirichlet form. There is no hidden axiomatization and no circular dependency.

\end{enumerate}

\begin{proof}
\input{proofCTheoremKineticEnergyIntrinsic}
\end{proof}

\end{theorem}

\begin{remark}[Resolving the Kinetic Energy Question]
\label{rem:resolvingKineticEnergy}

The original question was: ``Does the kinetic energy structure constitute a hidden third axiom?'' The answer, grounded in modern metric geometry, is \textbf{NO}. The kinetic energy is an intrinsic consequence of the Polish space metric measure structure (Axiom I).

Moreover, the minimal upper gradient approach is \textit{more general} than classical differential-geometric methods (which assume smooth manifolds). By working with metric Sobolev spaces, the framework achieves maximum generality while maintaining rigorous mathematical grounding.

The two-stage construction (Stage 1: pre-spectral form from metric structure; Stage 2: post-spectral verification using eigenfunctions) is workaround but rather a sophisticated and mathematically canonical approach to rigorously formalize the relationship between metric structure and spectral theory.

\end{remark}

