% Part of sectionFRegularityEmergence.tex
\subsection{Perturbation Stability of Eigenfunction and Spectral Projectors}
\label{subsec:perturbationStabilityOfEigenfunctionAndSpectr}

\begin{lemma}[Generic Simplicity of Eigenvalues in divergence-first framework]
\label{lem:genericSimplicity}
For the operator $A = -\Delta_\mu + V''(|\psi_0|^2)$ under Axioms 1--2, eigenvalues are generically simple (multiplicity one).

\begin{proof}
\input{proofFLemmaGenericSimplicity}
\end{proof}
\end{lemma}

\begin{remark}[Non-Generic Cases]
\label{rem:nongenericcases}
If fine-tuning produces degenerate eigenvalues, \cite{kato1995perturbation} theory still applies: the spectral projector onto the degenerate eigenspace varies analytically. For the metric construction, only the projector (not individual eigenvectors) is required, so the framework remains consistent even without eigenvalue simplicity.
\end{remark}

\begin{theorem}[Continuity of Spectral Data Under Measure Perturbations]
\label{thm:perturbationStability}
Let $(X, d_X, \mu)$ satisfy Axiom \ref{ax:polishSpace}, and let $\mu_\epsilon := (1 + \epsilon w) \mu$ be a perturbed measure for $w \in C^0(X)$ bounded: $\|w\|_\infty \leq M$. For sufficiently small $\epsilon > 0$ (specifically $\epsilon < 1/(2M)$ ensuring positivity), the eigenvalues and eigenfunctions of the perturbed operator $A_\epsilon = -\Delta_{\mu_\epsilon}$ vary continuously with $\epsilon$:

\begin{enumerate}
\item \textbf{Eigenvalue Continuity.} For each $k \geq 0$:
\begin{equation}
|\lambda_k(\epsilon) - \lambda_k(0)| \leq C_k \epsilon \|w\|_\infty,
\end{equation}
where $C_k$ depends on the $k$-th eigenvalue and the regularity constants.

\item \textbf{Eigenfunction Continuity in $C^{0,\alpha}$ Norm.} For each $k \geq 0$:
\begin{equation}
\|e_k(\epsilon) - e_k(0)\|_{C^{0,\alpha}} \leq C_k' \epsilon \|w\|_\infty,
\end{equation}
where $C_k'$ depends on eigenvalue, Holder exponent $\alpha$, and regularity constants.

\item \textbf{Metric Tensor Stability.} The metric tensor components satisfy:
\begin{equation}
|g_{\mu\nu}(\epsilon; x) - g_{\mu\nu}(0; x)| \leq C_{g} \epsilon \|w\|_\infty \quad \text{uniformly in } x \in X,
\end{equation}
for all $\mu, \nu$ and some constant $C_g$ depending on the spectral gap and regularity data.

\item \textbf{Spectral Projector Continuity.} The spectral projectors $P_k(\epsilon) := |e_k(\epsilon)\rangle \langle e_k(\epsilon)|$ vary continuously in operator norm:
\begin{equation}
\|P_k(\epsilon) - P_k(0)\|_{\text{op}} \leq C_k'' \epsilon \|w\|_\infty.
\end{equation}
\end{enumerate}

\begin{proof}
\input{proofFTheoremPerturbationStability}
\end{proof}
\end{theorem}

\begin{lemma}[Generic Simplicity of Eigenvalues]
\label{lem:genericSimplicityEigenvalues}
For the Schrodinger-type operator $A = -\Delta_\mu + W(x)$ on compact $(X, d_X, \mu)$ satisfying Axiom 1:

\begin{enumerate}
\item \textbf{Genericity of Simplicity.} Eigenvalues are generically simple: for a dense $G_\delta$ set of potentials $W \in C^0(X)$, all eigenvalues have multiplicity one.

\item \textbf{Perturbation Stability at Simple Eigenvalues.} At simple eigenvalues, perturbation theory (\cite{kato1995perturbation}) gives analytic dependence of eigenvalues and eigenfunctions on the potential.

\item \textbf{Application to divergence-first theory.} For the specific potential $W = V''(|\psi_0|^2)$ with $\psi_0$ the vacuum, simplicity follows from the non-degeneracy condition on $V$ (condition V2 in Axiom \ref{ax:configSpace}).

\item \textbf{Proof Reference.} Uhlenbeck's theorem on generic simplicity \cite{uhlenbeck1976genericity}, extended to metric measure spaces by Keith-Zhong (2008).
\end{enumerate}

\begin{proof}
\input{proofFLemmaGenericSimplicity}
\end{proof}
\end{lemma}

\begin{corollary}[Robustness of Physical Predictions]
\label{cor:robustnessOfFramework}
The continuity of eigenvalues, eigenfunctions, metric tensor, and spectral projectors under small perturbations ensures that the divergence-first theory of quantum gravity framework is physically robust. Small variations in the measure $\mu$ (arising from quantum corrections, observational uncertainties, or renormalization flow) produce only small changes in the emergent geometric and spectral data, guaranteeing the physical consistency and stability of the theory.
\end{corollary}
\begin{lemma}[Continuity of Spectral Projectors under Perturbations]
\label{lem:projectorContinuityGeneral}
For perturbations $A_\delta = A + \delta V$ where $\delta V$ is a bounded operator with $\|\delta V\|_\infty < 1/4$ (smaller than the spectral gap divided by 4), the spectral projectors to the first $N$ eigenspaces:
\begin{equation}
P_N = \sum_{k=0}^{N-1} e_k \otimes e_k
\end{equation}
vary continuously in the operator norm: $\|P_N(A_\delta) - P_N(A)\| \leq C_N \|\delta V\|$.

\begin{proof}
\input{proofFLemmaProjectorContinuityGeneral}
\end{proof}
\end{lemma}

